---
title: "Adaptive thresholding"
author: "Mira Sohn"
output:
    html_document:
        code_folding: hide
        df_print: paged
        # toc: true
        # toc_float: true
        # toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,
                      message=FALSE,
                      cache.lazy=FALSE)
```

Last run: `r date()`

This workflow is designed to demonstrate image segmentation for Visium analysis 
using the [squidpy (Palla, Spitzer et al., 2022)](https://doi.org/10.1038/s41592-021-01358-2) package.
Refer to the following resources for technical details:

- [Squidpy documentation](https://squidpy.readthedocs.io/en/stable/index.html)
- [Tutorial: Analyze Visium fluorescence data](https://squidpy.readthedocs.io/en/stable/notebooks/tutorials/tutorial_visium_fluo.html)
- [Squidpy GitHub](https://github.com/scverse/squidpy)


```{r rpackages}
library(reticulate)
library(tidyverse)
use_condaenv(params[['conda_env']])
source('config/helpers.R')
```

```{python ppackages}
import numpy as np
import dask.array as da
import matplotlib.pyplot as plt
import squidpy as sq
# from skimage.filters import threshold_local
from dask_image.ndfilters import threshold_local
import xarray as xr
import os
```

```{python config}
# ------------------------------------------------------------------------------
# This chunk configures user-specified variables
# ------------------------------------------------------------------------------

# reticulate::repl_python()

# Convert `params` from R to Python
params = r.params

# Paths to input and output obj
input_obj = params['input_obj']
output_obj = params['output_obj']

# Specify the name of input image layer to be processed
lyr = params['lyr']
lyr_smth = f"{lyr}_smooth"

# Set parameters for adaptive thresholding
# - block_size: neighborhood size, ~5–15% of the image dimension, odd integer.
# - offset: a constant added to (or subtracted from) the computed threshold (set to 0 by default)
#           higher offset results in more stringent thresholding
block_size = params['block_size']
offset = int(params['offset'])

# Path to output directory
outdir = params['outdir']
```

# Loading input image

```{python loading_input}

# ------------------------------------------------------------------------------
# This chunk imports input image saved as zarr in the previous step
# ------------------------------------------------------------------------------

# Load the Zarr store into an ImageContainer
# Set lazy=True to enable Dask-backed lazy loading, which is memory-efficient for large images.
# You can also specify chunks for Dask if needed.
img = sq.im.ImageContainer.load(input_obj, lazy=True, chunks='auto')

print(f"The following file has been loaded: {input_obj}")
print(img)
```


# Adaptive thresholding {.tabset}

[Adaptive thresholding (aka local thresholding)](https://scikit-image.org/docs/0.25.x/auto_examples/applications/plot_thresholding_guide.html#local-thresholding) 
is applied to achieve more uniform thresholding results across the pixels,
using the `threshold_local` function from the `scikit-image` package. Threshold values are
calculated as the weighted mean of the local neighborhood minus an offset value.

```{python add_local_thresholding}

# --------------------------------------------------------------------------------
# This chunk generates images with adaptive thresholding
# --------------------------------------------------------------------------------

# If the `block_size` is None, set it to a half of the height. Ensure to have it being an add number. 
if block_size == "None":
    block_size = round(img[lyr].shape[0] / 2 + 1)

for ch in range(img[lyr].shape[3]):
    # Specify the name of layer for the new processed image
    new_layer = f"adaptive_channel_{ch}"
    # Retrieve an array corresponding to the channel
    arr = img[lyr_smth].data[:, :, :, ch]
    # Compute adaptive threshold
    threshold_value = threshold_local(arr, block_size=arr.chunksize)

    # Binarize the array
    binary = arr > threshold_value
    # Add the binary array to the `ImageContainer` obj
    img.add_img(binary, layer=new_layer)

# Prep a list consisting of channel corresponding to adaptive-thresholded images
bn_lyrs = [c for c in list(img) if "adaptive_channel" in c]

# Create a list storing all dask arrays generated by adaptive thresholding
bn_arrays = [img[l].data for l in bn_lyrs]

# Stack the arrays along the last axis
stacked_array = da.stack(bn_arrays, axis=-1)
stacked_array = stacked_array.reshape(img[lyr].shape)

# Add the stacked array to the ImageContainer obj
lyr_bn = f"{lyr}_adaptive"
try:
    img[lyr_bn] = stacked_array
except:
    ValueError("Can't add the stacked array to ImageContainer. Wrong dimension!")

# Delete unnecessary layers
for l in bn_lyrs:
    del img[l]

print("ImageContainer updated:")
print(img)

```


```{r visualize_images, results='asis'}

# ------------------------------------------------------------------------------
# This chunk is used to prepare visualizing binary images
# ------------------------------------------------------------------------------

# Visualize binary images across the channels
for (i in 1:py$img[[py$lyr]][['shape']][[4]]) {
    # Convert into 0-indexing
    i <- i - 1
    cat("## Channel", i, "{.tabset}\n\n")
    # Command to show image
    t_deparsed_1 <- paste0("img.show('", py$lyr_bn, "', channel=", i, ", cmap='gray')")
    t_deparsed_2 <- paste0("plt.savefig('", py$outdir, "/adaptive_channel_", i,  ".png', dpi=1000)")
    t_deparsed_3 <- "plt.close()"
    t_deparsed <- c(t_deparsed_1, t_deparsed_2, t_deparsed_3)
    subchunkify(name=paste0("adaptive_", i), t_deparsed=t_deparsed)

    save_path <- paste0(py$outdir, "/adaptive_channel_", i, ".png")
    link_output(paste0("adaptive_channel_", i, ".png"))
}

```

# Saving the `ImageContainer` object

Output images are saved as the following object:

```{python save_zarr}
img.save(output_obj)
print(output_obj)
```


```{r session_info, collapse=FALSE}
sessionInfo()
```
