---
title: "Building ImageContainer object"
author: "Mira Sohn"
output:
    html_document:
        code_folding: hide
        df_print: paged
        # toc: true
        # toc_float: true
        # toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,
                      message=FALSE,
                      cache.lazy=FALSE)
```

Last run: `r date()`

This workflow is designed to demonstrate image segmentation for Visium analysis 
using the [squidpy (Palla, Spitzer et al., 2022)](https://doi.org/10.1038/s41592-021-01358-2) package.
Refer to the following resources for technical details:

- [Squidpy documentation](https://squidpy.readthedocs.io/en/stable/index.html)
- [Tutorial: Analyze Visium fluorescence data](https://squidpy.readthedocs.io/en/stable/notebooks/tutorials/tutorial_visium_fluo.html)
- [Squidpy GitHub](https://github.com/scverse/squidpy)


```{r rpackages}
library(reticulate)
library(tidyverse)
use_condaenv(params[['conda_env']])
source('config/helpers.R')
```

```{python ppackages}
import numpy as np
import matplotlib.pyplot as plt
import squidpy as sq
from matplotlib.colors import ListedColormap
from skimage.exposure import equalize_adapthist
import xarray as xr
import os
```

```{python configs}

# ------------------------------------------------------------------------------
# This chunk is used to specify variables for paths to input files/directories
# and parameter settings
# ------------------------------------------------------------------------------

# reticulate::repl_python()

# Convert `params` from R to Python
params = r.params

# Specify file paths to input and output files
outdir = params['outdir']
input_image = params['input_image']
output_obj = params['output_obj']

# Specify the name of input image layer to be processed
lyr = params['lyr']

# Specify the size of chunks for height and width
chunksize = int(params['chunksize'])

# Specify whether input images will be cropped
crop_images = params['crop_images']

# Specify coordinates to crop images
crop_height = float(params['crop_height'])
crop_width = float(params['crop_width'])
crop_size = float(params['crop_size'])
crop_scale = float(params['crop_scale'])
```

# Loading input images {.tabset}

We use `tif` image files converted from `vsi`. The following channels correspond to individual IFs.

- **Channel 0**: DAPI
- **Channel 1**: IBA1
- **Channel 2**: TDP43
- **Channel 3**: MAP2
- **Channel 4**: ALDH1L1

The following image object loaded:

```{python load_image}

# ------------------------------------------------------------------------------
# This chunk is used to load input image files
# ------------------------------------------------------------------------------

# Initialize ImageContainer
img = sq.im.ImageContainer(input_image, layer=lyr, lazy=True, chunks=chunksize)

# Crop input images if specified
if crop_images == "Y":
    print("Image cropped. Refer to the following dimension:", crop_coord)
    img = img.crop_corner(x=crop_width,
                          y=crop_height,
                          size=crop_size,
                          scale=crop_scale)
    print("ImageContainer obj updated:", img)

```

`ImageContainer` obj created using input image:

```{python print_status}
print(input_image)
print(img)
```


```{r vis_input, results='asis'}

# ------------------------------------------------------------------------------
# This chunk iteratively visualizes images before and after smoothing
# ------------------------------------------------------------------------------

# Visualize input images across the channels
for (i in 1:py$img[[py$lyr]][['shape']][[4]]) {
    # Convert into 0-indexing
    i <- i - 1
    cat("## Channel", i, "{.tabset}\n\n")
    # Command to show image
    t_deparsed_1 <- paste0("img.show('", py$lyr, "', channel=", i, ", cmap='gray')")
    t_deparsed_2 <- paste0("plt.savefig('", py$outdir, "/input_channel_", i,  ".png', dpi=1000)")
    t_deparsed <- c(t_deparsed_1, t_deparsed_2)
    subchunkify(name=paste0("input_", i), t_deparsed=t_deparsed)

    save_path <- paste0(py$outdir, "/input_channel_", i, ".png")
    link_output(paste0("input_channel_", i, ".png"))
}
```


# Saving the `ImageContainer` object

Output images are saved as the following object:

```{python save_zarr}
img.save(output_obj)
print(output_obj)
```

```{r session_info, collapse=FALSE}
sessionInfo()
```
