---
title: "Post processing with configurable cleaning"
author: "Mira Sohn"
output:
    html_document:
        code_folding: hide
        df_print: paged
        # toc: true
        # toc_float: true
        # toc_depth: 3
---

```{r setup, include=FALSE}
# Keep report rendering quiet so the HTML focuses on generated QC outputs.
knitr::opts_chunk$set(warning=FALSE,
                      message=FALSE,
                      cache.lazy=FALSE)
```

Last run: `r date()`

This workflow is designed to demonstrate image segmentation for Visium analysis
using the [squidpy (Palla, Spitzer et al., 2022)](https://doi.org/10.1038/s41592-021-01358-2) package.
Refer to the following resources for technical details:

- [Squidpy documentation](https://squidpy.readthedocs.io/en/stable/index.html)
- [Tutorial: Analyze Visium fluorescence data](https://squidpy.readthedocs.io/en/stable/notebooks/tutorials/tutorial_visium_fluo.html)
- [Squidpy GitHub](https://github.com/scverse/squidpy)


```{r rpackages}
# reticulate exposes Snakemake/R Markdown params to the Python chunks below.
library(reticulate)
library(tidyverse)
use_condaenv(params[['conda_env']])

# Helper functions render repeated per-channel Python visualizations in tabsets.
source('config/helpers.R')
```

```{python ppackages}
# Python handles the image-array work; R is used mainly for report rendering.
import numpy as np
import dask.array as da
import matplotlib.pyplot as plt
import squidpy as sq
from scipy.ndimage import generate_binary_structure

# Use scipy morphology on whole-channel NumPy masks after global parameter
# selection. This avoids choosing cleaning parameters independently per chunk.
from scipy.ndimage import binary_erosion as scipy_binary_erosion
from scipy.ndimage import binary_dilation as scipy_binary_dilation
from skimage.filters import threshold_otsu
from skimage.measure import label
from skimage.morphology import remove_small_objects
import pandas as pd
import os
```

```{python config}
# ------------------------------------------------------------------------------
# Parse Snakemake/R Markdown params, validate cleaning options, and resolve layer names.
# ------------------------------------------------------------------------------

# reticulate::repl_python()

# Convert `params` from R to Python
params = r.params


# Helper parsers keep Snakemake/R Markdown params consistent across old and
# new Snakefile versions.
def get_param(name, default=None):
    # Return a default when params are omitted or explicitly NULL-like.
    try:
        value = params[name]
    except Exception:
        return default
    if value is None:
        return default
    return value


def as_string(name, default=""):
    # Normalize R/CLI-style missing values before downstream type conversion.
    value = get_param(name, default)
    if value is None:
        return default
    value = str(value)
    if value in {"None", "NULL", "NA", "nan"}:
        return default
    return value


def as_int(name, default):
    # Convert numeric params safely after string normalization.
    value = as_string(name, str(default))
    if value == "":
        return int(default)
    return int(float(value))


def as_float(name, default):
    # Floats are used for quantiles and foreground-fraction thresholds.
    value = as_string(name, str(default))
    if value == "":
        return float(default)
    return float(value)


# Paths to the input Zarr, output Zarr, and machine-readable cleaning summary.
input_obj = as_string('input_obj')
output_obj = as_string('output_obj')
summary_out = as_string('summary_out')

# Base image layer and selected thresholding method. The method is resolved
# below to the concrete binary-mask layer used for cleaning.
lyr = as_string('lyr', 'image')
thresholding = as_string('thresholding', 'squidpy').lower()

# Specify the size of chunks for height and width
chunksize = as_int('chunksize', 5000)

# Path to output directory
outdir = as_string('outdir', os.path.dirname(output_obj))
if summary_out == "":
    summary_out = os.path.join(outdir, 'cleaning_summary.tsv')

# Cleaning options. Fixed mode preserves historical behavior; adaptive mode
# estimates per-channel settings; off mode passes the mask through unchanged.
cleaning_mode = as_string('cleaning_mode', 'fixed').lower()
morphology_mode = as_string('morphology_mode', 'fixed').lower()
min_size = as_int('minsize', 20)
adaptive_minsize_method = as_string('adaptive_minsize_method', 'log_otsu').lower()
adaptive_minsize_quantile = as_float('adaptive_minsize_quantile', 0.10)
adaptive_minsize_floor = as_int('adaptive_minsize_floor', min_size)
adaptive_minsize_ceiling = as_int('adaptive_minsize_ceiling', 500)
adaptive_min_components = as_int('adaptive_min_components', 25)
fixed_erosion_connectivity = as_int('fixed_erosion_connectivity', 2)
fixed_dilation_connectivity = as_int('fixed_dilation_connectivity', 1)
adaptive_foreground_fraction_min = as_float('adaptive_foreground_fraction_min', 0.001)
adaptive_foreground_fraction_max = as_float('adaptive_foreground_fraction_max', 0.50)

# Validate config early so failures point to the configuration problem rather
# than an obscure downstream array operation.
valid_cleaning_modes = {'fixed', 'adaptive', 'off'}
valid_morphology_modes = {'fixed', 'adaptive', 'off'}
valid_minsize_methods = {'log_otsu', 'quantile'}

if cleaning_mode not in valid_cleaning_modes:
    raise ValueError(f"cleaning_mode must be one of {sorted(valid_cleaning_modes)}")
if morphology_mode not in valid_morphology_modes:
    raise ValueError(f"morphology_mode must be one of {sorted(valid_morphology_modes)}")
if adaptive_minsize_method not in valid_minsize_methods:
    raise ValueError(f"adaptive_minsize_method must be one of {sorted(valid_minsize_methods)}")
if not 0 <= adaptive_minsize_quantile <= 1:
    raise ValueError("adaptive_minsize_quantile must be between 0 and 1")
if adaptive_minsize_floor < 1:
    raise ValueError("adaptive_minsize_floor must be >= 1")
if adaptive_minsize_ceiling < adaptive_minsize_floor:
    raise ValueError("adaptive_minsize_ceiling must be >= adaptive_minsize_floor")
if fixed_erosion_connectivity not in {1, 2} or fixed_dilation_connectivity not in {1, 2}:
    raise ValueError("fixed morphology connectivity must be 1 or 2 for 2D masks")

# Resolve the selected thresholding option to the layer created upstream.
# `watershed_segmentation.Rmd` consumes the final `image_removal` layer.
if thresholding in ['adaptive', 'otsu']:
    lyr_thr = f"{lyr}_{thresholding}"
elif thresholding in ['squidpy', 'sqbn']:
    lyr_thr = f"{lyr}_sqbn"
else:
    raise ValueError('thresholding must be adaptive, otsu, squidpy, sqbn, or stop before post_processing')

# Standard output layer names preserve downstream compatibility and existing
# visual-review expectations.
lyr_erosion = f"{lyr}_erosion"
lyr_dilation = f"{lyr}_dilation"
lyr_removal = f"{lyr}_removal"

sample_name = os.path.basename(os.path.normpath(outdir))

# Connectivity 2 includes diagonal neighbors; connectivity 1 uses orthogonal
# neighbors only. These defaults match the original fixed-cleaning workflow.
structure_fixed_erosion = generate_binary_structure(rank=2, connectivity=fixed_erosion_connectivity)
structure_fixed_dilation = generate_binary_structure(rank=2, connectivity=fixed_dilation_connectivity)
structure_adaptive = generate_binary_structure(rank=2, connectivity=1)

print("Post-processing configuration:")
print("- thresholding layer:", lyr_thr)
print("- cleaning_mode:", cleaning_mode)
print("- morphology_mode:", morphology_mode)
print("- fixed min_size:", min_size)
print("- summary_out:", summary_out)
```

# Loading input image

```{python loading_input}

# ------------------------------------------------------------------------------
# Load the upstream ImageContainer and confirm the selected mask layer exists.
# ------------------------------------------------------------------------------

# Load the Zarr store lazily so large images stay Dask-backed until a whole
# channel must be materialized for global component statistics.
# The configured chunksize is still used when the object is loaded and saved.
img = sq.im.ImageContainer.load(input_obj, lazy=True, chunks=chunksize)

# Fail before expensive work if the requested upstream mask does not exist.
if lyr_thr not in list(img):
    raise ValueError(f"Required thresholding layer '{lyr_thr}' not found. Available layers: {list(img)}")

print(f"The following file has been loaded: {input_obj}")
print(img)
```

# Cleaning masks

The selected cleaning mode is `r py$cleaning_mode`. The selected thresholding
layer is `r py$lyr_thr`.

```{python clean_masks}

# ------------------------------------------------------------------------------
# This chunk applies fixed, adaptive, or pass-through foreground mask cleaning.
# Adaptive parameters are estimated once per channel before cleaning, rather than
# separately inside each Dask chunk.
# ------------------------------------------------------------------------------


def compute_bool_channel(layer, channel):
    # Pull one full channel into memory to apply one consistent set of
    # cleaning parameters across the channel.
    arr = img[layer].data[:, :, 0, channel]
    if hasattr(arr, 'compute'):
        arr = arr.compute()
    return np.asarray(arr).astype(bool)


def channel_chunks(mask):
    # Rechunk saved output masks without exceeding spatial dimensions.
    return (min(chunksize, mask.shape[0]), min(chunksize, mask.shape[1]))


def to_dask_channel(mask):
    # Convert computed NumPy masks back to Dask arrays for ImageContainer storage.
    return da.from_array(mask.astype(bool), chunks=channel_chunks(mask))


def stack_channels(channels):
    # Restore Squidpy's expected image shape: height, width, z, channel.
    stacked = da.stack(channels, axis=-1)
    stacked = stacked[:, :, None, :]
    return stacked.rechunk((min(chunksize, stacked.shape[0]),
                            min(chunksize, stacked.shape[1]),
                            1,
                            1))


def component_summary(mask):
    # Connected-component areas drive adaptive min-size estimation and provide
    # before/after QC metrics for the summary TSV.
    labeled = label(mask)
    areas = np.bincount(labeled.ravel())[1:]
    foreground_pixels = int(mask.sum())
    total_pixels = int(mask.size)
    foreground_fraction = foreground_pixels / total_pixels if total_pixels else 0.0

    if len(areas) == 0:
        # Empty masks are valid; preserve NA-like values in the summary table.
        quantiles = [np.nan, np.nan, np.nan]
        min_area = np.nan
        max_area = np.nan
        mean_area = np.nan
    else:
        quantiles = np.quantile(areas, [0.10, 0.50, 0.90])
        min_area = int(np.min(areas))
        max_area = int(np.max(areas))
        mean_area = float(np.mean(areas))

    summary = {
        'foreground_pixels': foreground_pixels,
        'total_pixels': total_pixels,
        'foreground_fraction': float(foreground_fraction),
        'object_count': int(len(areas)),
        'area_min': min_area,
        'area_q10': float(quantiles[0]) if not np.isnan(quantiles[0]) else np.nan,
        'area_median': float(quantiles[1]) if not np.isnan(quantiles[1]) else np.nan,
        'area_q90': float(quantiles[2]) if not np.isnan(quantiles[2]) else np.nan,
        'area_max': max_area,
        'area_mean': mean_area,
    }
    return labeled, areas, summary


def bounded_min_size(value):
    # Clamp adaptive estimates so unusual channels cannot choose extremes.
    value = int(np.ceil(value))
    return min(max(value, adaptive_minsize_floor), adaptive_minsize_ceiling)


def estimate_adaptive_min_size(areas):
    # Estimate one min_size per channel from the global area distribution.
    # The log-Otsu path separates tiny speckles from larger objects.
    fallback_reason = ''

    if len(areas) < adaptive_min_components:
        # Too few objects makes area-distribution estimates unstable.
        return adaptive_minsize_floor, adaptive_minsize_floor, 'floor', 'too_few_components'

    if adaptive_minsize_method == 'quantile':
        # Quantile mode is simpler and useful when component areas are unimodal.
        raw = np.quantile(areas, adaptive_minsize_quantile)
        return bounded_min_size(raw), int(np.ceil(raw)), 'quantile', fallback_reason

    try:
        log_areas = np.log1p(areas.astype(float))
        if len(np.unique(log_areas)) < 2:
            raise ValueError('component areas have fewer than two unique values')
        raw = np.expm1(threshold_otsu(log_areas))
        return bounded_min_size(raw), int(np.ceil(raw)), 'log_otsu', fallback_reason
    except Exception as exc:
        raw = np.quantile(areas, adaptive_minsize_quantile)
        fallback_reason = f'log_otsu_failed_quantile_fallback: {exc}'
        return bounded_min_size(raw), int(np.ceil(raw)), 'quantile', fallback_reason


def choose_morphology_action(pre_summary, selected_min_size):
    # Decide whether morphology should run after min_size selection.
    # The action string is recorded in the summary for reproducibility.
    if cleaning_mode == 'off' or morphology_mode == 'off':
        return 'off', False
    if cleaning_mode == 'fixed' or morphology_mode == 'fixed':
        return f'fixed_erosion_c{fixed_erosion_connectivity}_dilation_c{fixed_dilation_connectivity}', True

    if pre_summary['object_count'] == 0:
        # No foreground means there is nothing morphology can improve.
        return 'adaptive_skipped_empty', False
    if pre_summary['foreground_fraction'] < adaptive_foreground_fraction_min:
        # Sparse masks are easy to over-erode, so keep them unchanged.
        return 'adaptive_skipped_sparse', False
    if not np.isnan(pre_summary['area_median']) and pre_summary['area_median'] <= selected_min_size * 2:
        # If typical components are close to the removal cutoff, morphology may
        # remove plausible foreground objects rather than just noise.
        return 'adaptive_skipped_small_components', False
    if pre_summary['foreground_fraction'] > adaptive_foreground_fraction_max:
        return 'adaptive_opening_c1_dense_mask', True
    return 'adaptive_opening_c1', True


def apply_morphology(mask, action):
    # Apply the selected opening operation. Skipped actions return the original
    # mask for both intermediate layers.
    if action.startswith('fixed'):
        eroded = scipy_binary_erosion(mask, structure=structure_fixed_erosion)
        dilated = scipy_binary_dilation(eroded, structure=structure_fixed_dilation)
        return eroded, dilated
    if action.startswith('adaptive_opening'):
        eroded = scipy_binary_erosion(mask, structure=structure_adaptive)
        dilated = scipy_binary_dilation(eroded, structure=structure_adaptive)
        return eroded, dilated
    return mask, mask


def prefixed_stats(prefix, stats):
    return {f'{prefix}_{key}': value for key, value in stats.items()}


# Accumulate one 2D mask per channel, then stack once at the end. This keeps
# the ImageContainer update path consistent across cleaning modes.
num_channels = img[lyr].shape[3]
erosion_arrays = []
dilation_arrays = []
removal_arrays = []
summary_rows = []
any_morphology_applied = False

for ch in range(num_channels):
    # Each channel is cleaned independently, but parameters are estimated
    # globally from that channel's full mask.
    source_mask = compute_bool_channel(lyr_thr, ch)
    _, source_areas, source_summary = component_summary(source_mask)

    if cleaning_mode == 'off':
        # Pass-through mode supports reviewing raw thresholding output and
        # isolating effects of later watershed/merge steps.
        selected_min_size = 0
        raw_min_size = 0
        min_size_method_used = 'off'
        fallback_reason = ''
        morphology_action = 'off'
        morphology_applied = False
        eroded_mask = source_mask
        dilated_mask = source_mask
        final_mask = source_mask
    else:
        if cleaning_mode == 'fixed':
            # Fixed mode should match the original post_processing.Rmd behavior.
            selected_min_size = min_size
            raw_min_size = min_size
            min_size_method_used = 'fixed'
            fallback_reason = ''
        else:
            selected_min_size, raw_min_size, min_size_method_used, fallback_reason = estimate_adaptive_min_size(source_areas)

        morphology_action, morphology_applied = choose_morphology_action(source_summary, selected_min_size)

        # Morphology, when enabled, runs before small-object removal.
        eroded_mask, dilated_mask = apply_morphology(source_mask, morphology_action)
        any_morphology_applied = any_morphology_applied or morphology_applied

        removal_input = dilated_mask if morphology_applied else source_mask

        # remove_small_objects expects labeled components; convert back to a
        # binary mask after removing components below the selected size.
        removal_labels = label(removal_input)
        final_mask = remove_small_objects(removal_labels, min_size=selected_min_size) > 0

    _, final_areas, final_summary = component_summary(final_mask)

    erosion_arrays.append(to_dask_channel(eroded_mask))
    dilation_arrays.append(to_dask_channel(dilated_mask))
    removal_arrays.append(to_dask_channel(final_mask))

    # Record chosen parameters and before/after statistics so this step can be
    # audited without reopening large Zarr outputs.
    row = {
        'sample': sample_name,
        'outdir': outdir,
        'channel': ch,
        'threshold_layer': lyr_thr,
        'cleaning_mode': cleaning_mode,
        'morphology_mode': morphology_mode,
        'morphology_action': morphology_action,
        'morphology_applied': morphology_applied,
        'selected_min_size': selected_min_size,
        'raw_min_size': raw_min_size,
        'min_size_method_used': min_size_method_used,
        'fallback_reason': fallback_reason,
        'fixed_erosion_connectivity': fixed_erosion_connectivity,
        'fixed_dilation_connectivity': fixed_dilation_connectivity,
        'adaptive_minsize_method': adaptive_minsize_method,
        'adaptive_minsize_quantile': adaptive_minsize_quantile,
        'adaptive_minsize_floor': adaptive_minsize_floor,
        'adaptive_minsize_ceiling': adaptive_minsize_ceiling,
        'adaptive_min_components': adaptive_min_components,
        'adaptive_foreground_fraction_min': adaptive_foreground_fraction_min,
        'adaptive_foreground_fraction_max': adaptive_foreground_fraction_max,
    }
    row.update(prefixed_stats('before', source_summary))
    row.update(prefixed_stats('after', final_summary))
    summary_rows.append(row)

# Keep morphology review layers only when morphology was actually applied.
if any_morphology_applied:
    img[lyr_erosion] = stack_channels(erosion_arrays)
    img[lyr_dilation] = stack_channels(dilation_arrays)

# The final cleaned mask must always be available under this layer name because
# downstream watershed segmentation reads `image_removal`.
img[lyr_removal] = stack_channels(removal_arrays)

has_erosion_layer = lyr_erosion in list(img)
has_dilation_layer = lyr_dilation in list(img)

# Write a small TSV sidecar for Snakemake's declared summary output.
summary_df = pd.DataFrame(summary_rows)
summary_dir = os.path.dirname(summary_out)
if summary_dir:
    os.makedirs(summary_dir, exist_ok=True)
summary_df.to_csv(summary_out, sep='\t', index=False)

print("ImageContainer updated:")
print(img)
print("Cleaning summary written:", summary_out)
print(summary_df)
```

# Cleaning summary

```{r cleaning_summary, results='asis'}
# Render the machine-readable TSV into the HTML report for quick inspection.
readr::read_tsv(py$summary_out, show_col_types=FALSE) %>%
    knitr::kable()
```

# Erosion {.tabset}

Erosion is displayed when morphology was applied to at least one channel.

```{r visualize_erosion, results='asis'}

# ------------------------------------------------------------------------------
# Render erosion review images only when an erosion layer was generated.
# ------------------------------------------------------------------------------

if (isTRUE(py$has_erosion_layer)) {
    # Python uses 0-indexed channels, so the R loop converts from 1-indexing.
    for (i in 1:py$img[[py$lyr]][['shape']][[4]]) {
        # Convert R's 1-indexed loop value to the Python channel index.
        i <- i - 1
        cat("## Channel", i, "{.tabset}\n\n")
        # Build a Python plotting subchunk for this channel and save a PNG.
        t_deparsed_1 <- paste0("img.show('", py$lyr_erosion, "', channel=", i, ", cmap='gray')")
        t_deparsed_2 <- paste0("plt.savefig('", py$outdir, "/erosion_channel_", i,  ".png', dpi=1000)")
        t_deparsed_3 <- "plt.close()"
        t_deparsed <- c(t_deparsed_1, t_deparsed_2, t_deparsed_3)
        subchunkify(name=paste0("erosion_", i), t_deparsed=t_deparsed)

        link_output(paste0("erosion_channel_", i, ".png"))
    }
} else {
    cat("Morphology was not applied; no erosion layer was created.\n\n")
}

```

# Dilation {.tabset}

Dilation is displayed when morphology was applied to at least one channel.

```{r visualize_dilation, results='asis'}

# ------------------------------------------------------------------------------
# Render dilation review images only when a dilation layer was generated.
# ------------------------------------------------------------------------------

if (isTRUE(py$has_dilation_layer)) {
    # Python uses 0-indexed channels, so the R loop converts from 1-indexing.
    for (i in 1:py$img[[py$lyr]][['shape']][[4]]) {
        # Convert R's 1-indexed loop value to the Python channel index.
        i <- i - 1
        cat("## Channel", i, "{.tabset}\n\n")
        # Build a Python plotting subchunk for this channel and save a PNG.
        t_deparsed_1 <- paste0("img.show('", py$lyr_dilation, "', channel=", i, ", cmap='gray')")
        t_deparsed_2 <- paste0("plt.savefig('", py$outdir, "/dilation_channel_", i,  ".png', dpi=1000)")
        t_deparsed_3 <- "plt.close()"
        t_deparsed <- c(t_deparsed_1, t_deparsed_2, t_deparsed_3)
        subchunkify(name=paste0("dilation_", i), t_deparsed=t_deparsed)

        link_output(paste0("dilation_channel_", i, ".png"))
    }
} else {
    cat("Morphology was not applied; no dilation layer was created.\n\n")
}

```

# Removal of remaining small foreground objects {.tabset}

The final binary mask is saved as `r py$lyr_removal`.

```{r visualize_removal, results='asis'}

# Always visualize the final mask because downstream steps depend on it.

# ------------------------------------------------------------------------------
# Render the final cleaned mask layer for every channel.
# ------------------------------------------------------------------------------

for (i in 1:py$img[[py$lyr]][['shape']][[4]]) {
    # Convert R's 1-indexed loop value to the Python channel index.
    i <- i - 1
    cat("## Channel", i, "{.tabset}\n\n")
    # Build a Python plotting subchunk for this channel and save a PNG.
    t_deparsed_1 <- paste0("img.show('", py$lyr_removal, "', channel=", i, ", cmap='gray')")
    t_deparsed_2 <- paste0("plt.savefig('", py$outdir, "/removal_channel_", i,  ".png', dpi=1000)")
    t_deparsed_3 <- "plt.close()"
    t_deparsed <- c(t_deparsed_1, t_deparsed_2, t_deparsed_3)
    subchunkify(name=paste0("removal_", i), t_deparsed=t_deparsed)

    link_output(paste0("removal_channel_", i, ".png"))
    cat('\n\n')
}

```

# Cleaning layers in the `ImageContainer` object

```{python clean_layers}

# ------------------------------------------------------------------------------
# Remove intermediate layers while preserving downstream and review layers.
# ------------------------------------------------------------------------------

# Delete unnecessary layers. Keep the selected thresholding layer, final removal
# layer, and dilation layer when it exists for visual review.
layers_del = [
    lyr_erosion,
    'image_adaptive',
    'image_otsu',
    'image_sqbn',
    'image_smooth']
layers_keep = [lyr_removal, lyr_dilation, lyr_thr]
layers_del = [l for l in layers_del if l in list(img) and l not in layers_keep]

print("The following layers are deleted: ", layers_del)
for l in layers_del:
    del img[l]

```

`ImageContainer` updated:

```{r final_img}
# Print the final layer inventory after cleanup and before saving.
print(py$img)
```

# Saving the `ImageContainer` object

Output images are saved as the following object:

```{python save_zarr}
# Persist the updated ImageContainer for watershed segmentation and channel merge.
img.save(output_obj)
print(output_obj)
```


```{r session_info, collapse=FALSE}
# Capture package versions in the rendered report for reproducibility.
sessionInfo()
```
