---
title: "Smoothing"
author: "Mira Sohn"
output:
    html_document:
        code_folding: hide
        df_print: paged
        # toc: true
        # toc_float: true
        # toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,
                      message=FALSE,
                      cache.lazy=FALSE)
```

Last run: `r date()`

This workflow is designed to demonstrate image segmentation for Visium analysis 
using the [squidpy (Palla, Spitzer et al., 2022)](https://doi.org/10.1038/s41592-021-01358-2) package.
Refer to the following resources for technical details:

- [Squidpy documentation](https://squidpy.readthedocs.io/en/stable/index.html)
- [Tutorial: Analyze Visium fluorescence data](https://squidpy.readthedocs.io/en/stable/notebooks/tutorials/tutorial_visium_fluo.html)
- [Squidpy GitHub](https://github.com/scverse/squidpy)


```{r rpackages}
library(reticulate)
library(tidyverse)
use_condaenv(params[['conda_env']])
source('config/helpers.R')
```

```{python ppackages}
import numpy as np
import matplotlib.pyplot as plt
import squidpy as sq
import dask.array as da
from skimage.filters import threshold_local
import xarray as xr
import os
```

```{python config}
# ------------------------------------------------------------------------------
# This chunk configures user-specified variables
# ------------------------------------------------------------------------------

# reticulate::repl_python()

# Convert `params` from R to Python
params = r.params

# Paths to input and output obj
input_obj = params['input_obj']
output_obj = params['output_obj']

# Specify signal values to be tested for Gaussian smoothing
# Sigma represents the variance for normal distribution
gaussian_sigma = int(params['gaussian_sigma'])

# Specify the name of input image layer to be processed
lyr = params['lyr']

# Specify normalization method that will be used for smoothing
norm_method = params['norm_method']

# Specify the size of chunks for height and width
chunksize = int(params['chunksize'])

# Path to output directory
outdir = params['outdir']
```

# Loading input image

```{python loading_input}

# ------------------------------------------------------------------------------
# This chunk imports input image saved as zarr in the previous step
# ------------------------------------------------------------------------------

# Load the Zarr store into an ImageContainer
# Set lazy=True to enable Dask-backed lazy loading, which is memory-efficient for large images.
# You can also specify chunks for Dask if needed.
img = sq.im.ImageContainer.load(input_obj, lazy=True, chunks=chunksize)

print(f"The following file has been loaded: {input_obj}")
print(img)
```

# Smoothing {.tabset}

Smoothing is required to reduce noises before image masking. The current workflow uses 
[Gaussian smoothing](https://en.wikipedia.org/wiki/Gaussian_blur) by calling 
the `squidpy.im.process(..., method="smooth")`, which implements the `skimage.filters.gaussian` function 
in the `scikit-image` package in python.

```{python smoothing}

# --------------------------------------------------------------------------------
# This chunk runs Gaussian smoothing
# --------------------------------------------------------------------------------

# Specify input layer for smoothing
input_layer = lyr
if norm_method != "raw":
    input_layer = f"{lyr}_{norm_method}"

# Specify the name of layer containing Gaussian smoothing
lyr_smth = f"{lyr}_smooth"

# Run smoothing
sq.im.process(img,
              layer=input_layer,
              method="smooth",
              layer_added=lyr_smth,
              sigma=gaussian_sigma,
              lazy=True,
              chunks=chunksize)

# Clean unnecessary layers
layers_del = [f"{lyr}_{l}" for l in ['clahe', 'lognorm', 'percnorm'] if l != norm_method]
for l in layers_del:
    if l in list(img):
        del img[l]
```
`ImageContainer` updated:

```{r smoothened_obj}
py$img
```

```{r visualize_images, results='asis'}

# ------------------------------------------------------------------------------
# This chunk is used to prepare visualizing smoothened images
# ------------------------------------------------------------------------------

# Visualize input images across the channels
for (i in 1:py$img[[py$lyr]][['shape']][[4]]) {
    # Convert into 0-indexing
    i <- i - 1
    cat("## Channel", i, "{.tabset}\n\n")
    # Command to show image
    t_deparsed_1 <- paste0("img.show('", py$lyr_smth, "', channel=", i, ", cmap='gray')")
    t_deparsed_2 <- paste0("plt.savefig('", py$outdir, "/smooth_channel_", i,  ".png', dpi=1000)")
    t_deparsed_3 <- "plt.close()"
    t_deparsed <- c(t_deparsed_1, t_deparsed_2, t_deparsed_3)
    subchunkify(name=paste0("smooth_", i), t_deparsed=t_deparsed)
    
    link_output(paste0("smooth_channel_", i, ".png"))
    cat('\n\n')
}
```

# Saving the `ImageContainer` object

Output images are saved as the following object:

```{python save_zarr}
img.save(output_obj)
print(output_obj)
```

```{r session_info, collapse=FALSE}
sessionInfo()
```
