---
title: "Segmentation demo using squidpy"
author: "Mira Sohn"
output:
    html_document:
        code_folding: hide
        df_print: paged
        # toc: true
        # toc_float: true
        # toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,
                      message=FALSE,
                      cache.lazy=FALSE)
```

Last update: `r date()`

This workflow is designed to demonstrate image segmentation for Visium analysis 
using the [squidpy (Palla, Spitzer et al., 2022)](https://doi.org/10.1038/s41592-021-01358-2) package.
Refer to the following resources for technical details:

- [squidpy documentation](https://squidpy.readthedocs.io/en/stable/index.html)
- [squidpy GitHub](https://github.com/scverse/squidpy)

```{r rpackages}
library(reticulate)
library(tidyverse)
use_condaenv('../env')
```

```{python ppackages}
import numpy as np
import matplotlib.pyplot as plt
import squidpy as sq
import os
```

```{python configs}

# ------------------------------------------------------------------------------
# This chunk is used to specify variables for paths to input files/directories
# and parameter settings
# ------------------------------------------------------------------------------

# reticulate::repl_python()

# Specify argument setting for printing images
channel_config = {
    'Merged' : {'ch' : None, 'cm' : None},
    'Channel 0' : {'ch' : 0, 'cm' : 'gray'},
    'Channel 1' : {'ch' : 1, 'cm' : 'gray'},
    'Channel 2' : {'ch' : 2, 'cm' : 'gray'},
}

# Specify the name of input image layer to be processed
lyr = 'image'
```

# Loading demo images {.tabset}

We use demo images provided by the `squidpy` package. The image is loaded by calling 
`sq.datasets.visium_fluo_image_crop()`.

The following image object loaded:

```{python load_image}

# ------------------------------------------------------------------------------
# This chunk is used to load input image files
# ------------------------------------------------------------------------------

# Load fluorescence tissue images
img = sq.datasets.visium_fluo_image_crop()

print(img)
```

## All

```{python print_all_images, fig.height=12, fig.width=12}

# ------------------------------------------------------------------------------
# This chunk prints all input images at once
# e.g.
# fig, axes = plt.subplots(nrows=2, ncols=2)
# img.show("image", ax=axes[0, 0])
# _ = axes[0, 0].set_title("Merged")
# img.show("image", channel=0, ax=axes[0, 1], cmap='gray')
# _ = axes[0, 1].set_title("Channel 0")
# img.show("image", channel=1, ax=axes[1, 0], cmap='gray')
# _ = axes[1, 0].set_title("Channel 1")
# img.show("image", channel=2, ax=axes[1, 1], cmap='gray')
# _ = axes[1, 1].set_title("Channel 2")
# ------------------------------------------------------------------------------

# Specify plotting arguments
rows = 2
cols = 2

# Initialize subplots
fig, axes = plt.subplots(nrows=rows, ncols=cols)

# Prep subplot dimensions
subplot_dims = [axes[i, j] for i in range(rows) for j in range(cols)]

# Print plots
for i in range(len(subplot_dims)):
    # Pull pre-specified variables
    key = list(channel_config.keys())[i]
    subdim = subplot_dims[i]
    ch = channel_config[key]['ch']
    cm = channel_config[key]['cm']
    # Print
    img.show(lyr,
        ax=subdim,
        channel=ch,
        cmap=cm
        )
    # Add corresponding title
    _ = subdim.set_title(key)
```

## Individuals {.tabset}

### Merged

```{python individual_images_merged}
img.show(lyr)
```

### Channel 0

```{python individual_images_c0}
img.show(lyr, channel=0, cmap='gray')
```

### Channel 1

```{python individual_images_c1}
img.show(lyr, channel=1, cmap='gray')
```

### Channel 2

```{python individual_images_c2}
img.show(lyr, channel=2, cmap='gray')
```

# Converting to grayscale

```{python to_grayscale, eval=FALSE}

# --------------------------------------------------------------------------------
# This chunk converts the images to grayscale
# --------------------------------------------------------------------------------

sq.im.process(img, layer=lyr, method="gray")

```

```{r session_info}
sessionInfo()
```
